---
title: "General extenders in Irish English"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  bookdown::html_document2: default
bibliography: bibliography.bib
link-citations: yes
---

# Introducion

This document shows an analysis of general extenders in Irish English based on the Irish component of the International Corpus of English (ICE). 

# Preparation

Install packages

```{r prep01, eval = F, message=F, warning=F}
install.packages("tidyverse")
install.packages("quanteda")
install.packages("readtext")
install.packages("plyr")
install.packages("Boruta")
install.packages("ggparty")
install.packages("party")
install.packages("partykit")
install.packages("partykit")
```

Load packages

Set options

Define paths to corpus and biodata

```{r prep02, message=F, warning=F}
# load packages
library(tidyverse)
library(quanteda)
library(readtext)
library(here)
# set options
options(stringsAsFactors = F)                           
options(scipen = 999) 
options(max.print=10000)
# define image directory
imageDirectory<-"images"   
# specify paths
cpath <- "D:\\Uni\\Korpora\\Original\\ICE Ireland version 1.2.2\\ICE-Ireland txt\\ICE spoken running txt"
biopath <- "D:\\Uni\\Korpora\\Metadata/BiodataIceIreland.txt"
```

# Preprocessing

Create vector of corpus files.

```{r genex0101, message=FALSE, warning=FALSE}
# load corpus files
cfiles = list.files(path = cpath, all.files = T,
                          full.names = T, recursive = T, ignore.case = T,
                          include.dirs = T)
# inspect files
length(cfiles); head(cfiles)
```

Load corpus 

```{r genex0102, message=FALSE, warning=FALSE}
# load corpus
cire <- readtext::readtext(cfiles) %>%
  dplyr::mutate(doc_id = stringr::str_remove_all(doc_id, " .*"),
                text = stringr::str_squish(text),
                text = stringr::str_trim(text, side = "both"))
# inspect data
str(cire);head(cire)
```

# Processing

Split corpus into speakers and speechunits.

```{r genex0103, message=FALSE, warning=FALSE}
# create vector with number of speech units
idx <- stringr::str_count(cire$text, "<S[12][AB]")
# create vector of files that matches the number of speech units
fnames <- rep(cire$doc_id, idx)
# extract speech units
sup <- stringr::str_split(stringr::str_replace_all(cire$text, "(<S[12][AB])", "~~~\\1"), "~~~") %>%
  unlist() %>%
  stringr::str_subset(".+")
# combine file names and speech units in a data frame
ciredf <- data.frame(fnames, sup)
# inspect
head(ciredf)
```

Adding speaker

```{r genex0104, message=FALSE, warning=FALSE}
ciredf %>%
  dplyr::mutate(speaker = stringr::str_remove_all(sup, ">.*")) %>%
  dplyr::mutate(speaker = stringr::str_remove_all(speaker, ".*\\$")) %>%
  dplyr::filter(nchar(speaker) < 3) -> ciredf
# inspect
head(ciredf)
```

Split further into speech units

```{r genex0105, message=FALSE, warning=FALSE}
# extract speech units
su <- ciredf$sup %>%
  stringr::str_remove_all("<S[12][AB]-[0-9]{3,3}\\$\\?{0,1}[A-Z]{1,2}\\?{0,1}>") %>%
  stringr::str_squish() %>%
  stringr::str_split("<\\#>")
# extract length of speech units per sup
idx2 <- sapply(su, function(x) length(x))
# create vector of file names matching number of su 
file <- rep(ciredf$fnames, idx2)
# create vector of speakers matching number of su
speaker <- rep(ciredf$speaker, idx2)
# combine file names and speech units in a data frame
ire <- data.frame(file, speaker, unlist(su)) %>%
  dplyr::rename(su = colnames(.)[3]) %>%
  dplyr::filter(nchar(su) > 1) %>%
  dplyr::mutate(su_clean = stringr::str_remove_all(su, "<&> .*? </&>"),
                su_clean = stringr::str_remove_all(su_clean, "<unclear> .*? </unclear>"),
                su_clean = tolower(su_clean),
                su_clean = stringr::str_remove_all(su_clean, "[^[:alpha:]\\' ]"),
                id = 1:nrow(.)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(wordsunit = stringr::str_count(su_clean, "\\w+"))
# save data to disc
write.table(ire, here::here("tables", "iceraw.txt"), 
            sep = "\t", row.names = F, col.names = T)
# inspect
head(ire, 10)
```

## Biodata

Load biodata.

```{r genex0106, message=FALSE, warning=FALSE}
# read in data
bio <- read.table(biopath, sep = "\t", header=TRUE)
# clean file for joining
bio <- bio %>%
  dplyr::rename(File = text.id,
                Speaker = spk.ref,
                FileSpeakerId = file.speaker.id,
                DateOfRecording = date,
                Gender = sex,
                Age = age,
                SpeechUnits = speech.unit.count,
                Words = word.count,
                ReligiousAffiliation = relig,
                Occupation = occupation,
                Education = ed.lev,
                ChildhoodCounty = prov,
                ResidenceType = reside,
                Region = zone) %>%
  dplyr::select(-orig.id, -subfile, -spk.id, -title, -X1st.lg, -other.lgs) %>%
  dplyr::mutate(texttype = stringr::str_remove_all(File, "-.*"))
# inspect biodata
head(bio)
```

Clean biodata.

```{r genex0107, message=FALSE, warning=FALSE}
# remove speakers below the age of 18
bio <- bio %>%
  dplyr::filter(Age != "0-18") %>%
  # recode age
  dplyr::mutate(Age = ifelse(Age == "34-41", "34-49",
         ifelse(Age == "42-49", "34-49", Age))) %>%
  # clean education
  dplyr::mutate(Education = str_replace_all(Education, ".*doctoral degree.*",
                                            "College")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*first degree.*", "College")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*non-degree tertiary qualification.*", "College")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*postgraduate qualification.*",
                                            "College")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*some tertiary education.*", 
                                            "College")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*primary education.*", 
                                            "NoCollege")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*secondary school qualification.*",
                                            "NoCollege")) %>%
  dplyr::mutate(Education = str_replace_all(Education, 
                                            ".*some secondary education.*",
                                            "NoCollege"))
# inspect data
table(bio$Education); table(bio$Age)
```

Recode ResidenceType

```{r genex0108, message=FALSE, warning=FALSE}
# recode ResidenceType
Urban <- c("cork (city)", "belfast", "dublin (city)")
Rural <- c("antrim", "armagh", "cork (county)", "down", 
           "dublin (county)", "kerry", "kildare", "limerick", 
           "(london)derry", "mayo", "meath", 
           "sligo", "tipperary", "tyrone", "wicklow")
bio <- bio %>%
  dplyr::mutate(ResidenceType = ifelse(ResidenceType %in% Urban, "Urban", 
                               ifelse(ResidenceType %in% Rural, "Rural", NA)))
# inspect data
table(bio$ResidenceType)
```

Recode ReligiousAffiliation

```{r genex0109, message=FALSE, warning=FALSE}
# recode ResidenceType
bio <- bio %>%
  dplyr::mutate(ReligiousAffiliation = ifelse(ReligiousAffiliation == "catholic",
                                              "Catholic",
                               ifelse(ReligiousAffiliation == "protestant",
                                      "Protestant", NA)))
# inspect data
table(bio$ReligiousAffiliation)
```


Recode Region

```{r genex0110, message=FALSE, warning=FALSE}
# recode ResidenceType
bio <- bio %>%
  dplyr::mutate(Region = dplyr::case_when(Region == "northern ireland" ~ "NorthernIreland",
                                          Region == "republic of ireland" ~ "RepublicOfIreland",  
                                          Region == "mixed between ni and roi" ~  "Interstate", 
                                          TRUE ~ as.character(NA)))
# inspect data
table(bio$Region)
```

Recode ChildhoodCounty

```{r genex0111, message=FALSE, warning=FALSE}
# recode ChildhoodCounty
North <- c("antrim", "armagh", "belfast", "(london)derry", "down", "tyrone")
South <- c("cork (city)",  "cork (county)", "dublin (city)", 
           "dublin (county)", "kerry", "kildare", "limerick", "mayo", "meath", 
           "sligo", "tipperary", "wicklow")
bio <- bio %>%
  dplyr::mutate(ChildhoodCounty = dplyr::case_when(ChildhoodCounty %in% North ~ "North",
                                                   ChildhoodCounty %in% South ~ "South", 
                                                   TRUE ~ as.character(NA)))
# inspect data
table(bio$ChildhoodCounty)
```

```{r genex0112, message=FALSE, warning=FALSE}
# recode occupation
# Academic, clerical, managerial professions
acmp <- c("accountant", "administrator", "apprentice barrister", 
          "artist typist", "author", "author journalist", "bank clerk", 
          "bank manager", "bank official", "bank official lecturer", 
          "banker", "barrister", "bbc production assistant",
          "bifhe tutor", "biochemist", "bishop", "broadcast journalist",
          "broadcast reporter", "broadcaster", "broadcaster cooking school owner",
          "broadcaster director",
          "broadcaster journalist", "broadcaster journalist writer", 
          "broadcaster producer",
          "building society executive", "businesswoman", "catholic priest", "ceo",
          "chief auditor of ehsesb", "chief executive", "civil servant",
          "classroom assistant",
          "clerical officer", "college administrator", "college president",
          "commission chairman", "company director", 
          "company director politician",
          "company director sports com", "company representative", 
          "computer technician",
          "consultant", "court clerk", "credit union official", "curate",
          "department store manager", "deputy editor", "designer", 
          "diplomat politician",
          "doctor", "domestic economist", "economist politician", 
          "employment coordinator",
          "engineer", "environmental health officer", "estate agent",
          "executive director unicef irl", "financial advisor", 
          "financier author",
          "football manager", "former gaa hurler student", "former town clerk",
          "fun park owner", "general practitioner", "graphic designer",
          "health employers official", "high court judge", 
          "hotelier sports commentator",
          "investment analyst", "it analyst", "journalist", 
          "journalist  broadcaster",
          "journalist author", "journalist broadcaster", 
          "journalist community worker",
          "journalist producer", "journalist writer", "judge", 
          "lecturer", "lecturer musician",
          "lecturer security analyst", "legal secretary", "letting agent",
          "lord mayor politician", "management consultant", "manager", 
          "manager social worker",
          "marketing", "marketing assistant", "marketing consultant", 
          "marketing manager",
          "medical consultant", "medical doctor", "medical journal editor",
          "methodist minister",
          "minister of religion", "musician", "musician singer", 
          "news reporter", "newscaster",
          "newsreader", "newsreporter", "pastor community leader", 
          "pg student", "physicist",
          "physiotherapist student", "police superintendent", "politician",
          "politician academic", "politician accountant", "politician architect",
          "politician barrister", "politician barrister accountant",
          "politician business executive", "politician company director",
          "politician economist", "politician executive", "politician historian",
          "politician lecturer consultant", "politician minister of religion",
          "politician psychiatrist", "politician publican", 
          "politician secretary",
          "politician solicitor", "politician teacher", 
          "politician trade union official",
          "practice manager", "presbyterian minister", 
          "presbyterian moderator", "priest",
          "priest journalist", "priest lecturer", "priest teacher", 
          "professor",
          "professor composer", "programme manager trocaire", 
          "property developer politician",
          "public relations officer", "publican", "radio presenter",
          "radio producer writer teacher", "registrar", 
          "rep for chemical company",
          "reporter", "research assistant", "researcher", "retired accountant",
          "retired bishop", "retired broadcaster writer", "retired civil servant",
          "retired general secretary gra", "retired presbyterian minister",
          "retired social worker", "sales manager", "secretary student",
          "senior counsel lecturer", "senior legal assistant", 
          "senior systems analyst",
          "singer", "solicitor", "solicitor politician", 
          "speech & language therapist",
          "sports commentator", "sports commentator teacher", "sports official",
          "sports reporter", "student", "student hockey player", 
          "student union officer",
          "surgeon", "systems analyst", "teacher", "teacher politician",
          "television journalist", "television painting expert", 
          "television presenter",
          "television producer ski instructor", "tourist board employee",
          "trade union official", "trade union representative", 
          "trainee solicitor", "travel advisor",
          "vp of primary school", "weather reporter", "writer", 
          "writer broadcaster")
# Skilled Manual Labour
sml <- c("athlete", "staff nurse", "waitress", "ward assistant", 
         "social worker",
         "socialworker", "building company employee", "bus driver", 
         "captain of fishing boat",
         "retail manageress", "shop employee", "retired social worker", 
         "care assistant",
         "community worker", "racing commentator", "receiver", 
         "receptionist", "farm manager",
         "farmer", "farmer politician", "politician stonemason", 
         "fire officer", "fisherman",
         "fishwife", "fitness instructor", "footballer", "former gaa footballer",
         "furniture store employee", "former road sweeper", 
         "horticulturist", "nanny",
         "nurse", "nurse student", "peat bog owner", "police officer")
# unclassifiable
NAN <- c("retired", "employed", "housewife", "na")
# reclassify
bio <- bio %>%
  dplyr::mutate(Occupation = dplyr::case_when(Occupation %in% acmp ~ "AcademicManagerialProfessions",
                                               Occupation %in% sml ~ "SkilledManualLabour", 
                                                   TRUE ~ as.character(NA)))
# inspect data
table(bio$Occupation)
```

AudienceSize

```{r genex0113, message=FALSE, warning=FALSE}
bio <- bio %>%
  dplyr::group_by(File) %>%
  dplyr::mutate(AudienceSize = length(table(Speaker))) %>%
  dplyr::mutate(AudienceSize = dplyr::case_when(AudienceSize <= 2 ~ "Dyad",
                                                AudienceSize >= 6 ~ "LargeGroup",
                                                TRUE ~ "SmallGroup"))
# inspect data
table(bio$AudienceSize)
```
ConversationType

```{r genex0114, message=FALSE, warning=FALSE}
bio <- bio %>%
  dplyr::group_by(File) %>%
  dplyr::mutate(ConversationType = sum(ifelse(table(Gender) > 0, 1, 0))) %>%
  dplyr::mutate(ConversationType = dplyr::case_when(ConversationType == 1 ~ "SameGender",
                                          TRUE ~ "MixedGender"))
# inspect data
table(bio$ConversationType)
```

# Combine Corpus and Biodata


```{r genex0115, message=FALSE, warning=FALSE}
colnames(ire) <- stringr::str_to_title(colnames(ire))
# merge
ire <- dplyr::left_join(ire, bio, by = c("File", "Speaker")) %>%
  dplyr::mutate(Su = stringr::str_squish(Su),
                Su_clean = stringr::str_squish(Su_clean)) %>%
  dplyr::select(-id)
# save data to disc
write.table(ire, here::here("tables", "ice.txt"), 
            sep = "\t", row.names = F, col.names = T)
# inspect
head(ire)
```


# Combine kwics with corpus and biodata

Extract general extenders

```{r genex0116, message=FALSE, warning=FALSE}
# define regular expression
regex <-  "(sort|sorts) of (stuff|shit)|(\\betc\bb|etcetera etcetera|etcetera|et cetera)|(and|or) (this|that){0,4} {0,1}(whatever|so on|so forth|all|something|everything|thing|things|stuff|shit) {0,1}(else|like){0,4} {0,1}(this|that|these|those){0,5}"
```



```{r genex0118, message=FALSE, warning=FALSE}
ge2 <- stringr::str_extract_all(ire$Su_clean, regex)
# extract number of genex for each su
ire$ngenex <- sapply(ge2, function(x) length(x))
# inspect
head(ire, 10)
```

split speech units so that all units containing 2 or more GEs are split into two speech unit 

```{r genex0119, message=FALSE, warning=FALSE}
splitsu <- ifelse(ire$ngenex > 1, stringr::str_split(str_replace(ire$Su_clean, (regex), "\\1~~~"), "~~~"), ire$Su_clean) %>%
  unlist()
```

chec with speech units still contain more than 1 GE

```{r genex0120, message=FALSE, warning=FALSE}
# now check which split units still contain more than one GE
ge3 <- stringr::str_extract_all(splitsu, regex)
# extract number of genex for each su
nge3 <- sapply(ge3, function(x) length(x))
```

repeat splitting until each speech unit only contains a maximum of 1 GE

```{r genex0121, message=FALSE, warning=FALSE}
# repeat split so that each su only conatins max 1 ge
splitsu <- ifelse(nge3 > 1, stringr::str_split(str_replace(splitsu, (regex), "\\1~~~"), "~~~"), splitsu) %>%
  unlist()
```

repeat all rows in the data so that the split speech units can be added again aligning correctly with the metadata of the speech unit

```{r genex0125, message=FALSE, warning=FALSE}
# repeat rows
reps <- ifelse(ire$ngenex == 0, 1, ire$ngenex)
ireb <- ire[rep(1:nrow(ire), reps), ]
# add splitsu
ireb$nsu <- splitsu
# extract general extenders
gxtb <- stringr::str_extract_all(ireb$nsu, regex)
# create vector of forms
gxs <- sapply(gxtb, function(x) ifelse(length(x) == 0, "0", x))
# add to table
ireb$gx <- gxs 
ireb <- ireb %>%
  # update ngx
  dplyr::mutate(ngenex = ifelse(gx == "0", 0, 1),
                gx = stringr::str_squish(gx),
                Speaker = stringr::str_remove_all(Speaker, "\\?")) %>%
  dplyr::filter(Speaker != "X",
                Speaker != "")
# save data to disc
write.table(ireb, here::here("tables", "ice_prefinal.txt"),
            sep = "\t", row.names = F, col.names = T)
# inspect 
head(ireb)
```

Check

```{r}
ireb %>%
  dplyr::filter(gx == "etcetera") %>%
  dplyr::select(nsu) %>%
  head(10)
```


## Overview{-}

 create table showing how often each GE occurred in the data



```{r genex0126, message=FALSE, warning=FALSE}
genex <- ge3 %>%
  unlist() %>%
  stringr::str_squish() %>%
  table() %>%
  as.tibble() %>%
  dplyr::rename(genex = 1) %>%
  dplyr::mutate(genex = stringr::str_squish(genex))
# inspect
genex

```

visualize overview


```{r genex0153, message=FALSE, warning=FALSE}
genex %>%
  dplyr::arrange(-n) %>%
  ggplot(aes(x = reorder(genex, -n), y = n, label = n)) +
  geom_bar(stat = "identity") +
  geom_text(vjust=-1.2, color = "gray30", size = 3.5) +
  labs(x = "", y = "Frequency") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  coord_cartesian(y = c(0, 200))
# save plot
ggsave(here::here("images", "ngenex.png"), unit = "cm", width = 25)
```


```{r}
clps <- genex %>%
  dplyr::filter(n < 5) %>%
  dplyr::pull(genex)
# inspect
clps
```



```{r genex0150, message=FALSE, warning=FALSE}
gx_genre <- ireb %>%
  dplyr::filter(gx != "0") %>%
  dplyr::mutate(gx = ifelse(gx %in% clps, "other", gx)) %>%
  dplyr::group_by(Age, texttype, gx) %>%
  dplyr::summarise(freq = n()) %>%
  dplyr::ungroup() %>%
  dplyr::rename(Genre = texttype,
                GeneralExtender = gx,
                Frequency = freq) %>%
  dplyr::arrange(-Frequency) %>%
  ggplot(aes(x = reorder(GeneralExtender, -Frequency), y = Frequency, label = Frequency)) +
  geom_bar(stat = "identity") +
  geom_text(vjust=-1.2, color = "gray30", size = 3.5) +
    facet_grid(vars(Genre)) +
  labs(x = "", y = "Frequency") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  coord_cartesian(y = c(0, 200))
# inspect
gx_genre
# save plot
ggsave(here::here("images", "ngx_tt.png"), unit = "cm", width = 25)
```


```{r genex0151, message=FALSE, warning=FALSE}
ireb %>%
  dplyr::filter(gx != "0") %>%
  dplyr::mutate(gx = ifelse(gx %in% clps, "other", gx)) %>%
  dplyr::group_by(Age, texttype, gx) %>%
  dplyr::summarise(freq = n()) %>%
  dplyr::ungroup() %>%
  dplyr::rename(Genre = texttype,
                GeneralExtender = gx,
                Frequency = freq) %>%
  dplyr::arrange(-Frequency) %>%
  ggplot(aes(x = reorder(GeneralExtender, -Frequency), y = Frequency, label = Frequency)) +
  geom_bar(stat = "identity") +
  geom_text(vjust=-1.2, color = "gray30", size = 3.5) +
    facet_grid(vars(Age)) +
  labs(x = "", y = "Frequency") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  coord_cartesian(y = c(0, 200))
# save plot
ggsave(here::here("images", "ngx_age.png"), unit = "cm", width = 25)
```

# Simplfy data for analysis

extract vector of rare GEs

```{r genex0128, message=FALSE, warning=FALSE}
freq_ges <- names(table(ireb$gx))[which(table(ireb$gx) > 40)] %>%
  stringr::str_squish()
# inspect
freq_ges
```


```{r genex0130, message=FALSE, warning=FALSE}
irec <- ireb %>%
  dplyr::mutate(ges = dplyr::case_when(gx == "0" ~ "0",
                                       gx %in% freq_ges ~ gx, 
                                       TRUE ~ "other"))
# save data to disc
write.table(irec, here::here("tables", "ice_prestats.txt"),
            sep = "\t", row.names = F, col.names = T)
# inspect
head(irec)
```





# Statistical Analysis

load packages


```{r}
library(Boruta)
library(ggparty)
library(party)
library(partykit)
library(cfa)
```



Prepare data

```{r genex0161, message=FALSE, warning=FALSE}
# create vector of variables to keep
keep <-c("File", "Speaker", "Id", "Wordsunit",  "Region", "DateOfRecording", 
         "Gender", "Age", "ChildhoodCounty", "ResidenceType", "Education", 
         "Occupation", "ReligiousAffiliation", "texttype", "AudienceSize", 
         "ConversationType", "ngenex",  "gx", "ges")
# simplify data
dstat <- irec %>%
  dplyr::select(keep) %>%
  dplyr::rename(nge = ngenex,
                ge = gx) %>%
  dplyr::mutate(texttype = dplyr::case_when(texttype == "S1A" ~ "Private",
                                            texttype == "S1B" ~ "Public",
                                            texttype == "S2A" ~ "Scripted",
                                            texttype == "S2B" ~ "Unscripted",
                                            TRUE ~ "other"),
                texttype = factor(texttype, 
                                  levels = c("Private", "Public", 
                                             "Unscripted",  "Scripted")),
                Age = factor(Age, levels = c("19-25", "26-33", "34-49", "50+"))) %>%
  dplyr::mutate_if(is.character, factor)
# inspect
head(dstat)
```

## Data Vis

```{r}
dstat %>%
  dplyr::select(Age, texttype, ges, nge, Wordsunit) %>%
  dplyr::group_by(Age, texttype, ges) %>%
  summarise(words = sum(Wordsunit),
            ns = sum(nge),
            freq = round(ns/words *1000, 3)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(ges != "0") %>%
  dplyr::rename(Variant = ges) %>%
  na.omit() %>%
  ggplot(aes(x = Age, y = freq, group = Variant, color = Variant)) +
  facet_wrap(~texttype, ncol = 1) +
  geom_line() +
  theme_bw() +
  labs(y = "Relative Frequency (mean per 1,000 words)")
ggsave(here::here("images", "ge_age_tt.png"), unit = "cm", width = 15)
```


```{r}
dstat %>%
  dplyr::select(Age, ge) %>%
  dplyr::mutate(Variant = dplyr::case_when(str_detect(ge, "stuff") ~ "stuff",
                                         str_detect(ge, "shit") ~ "shit",
                                         str_detect(ge, "thing") ~ "thing",
                                         TRUE ~ "other")) %>%
  dplyr::filter(Variant != "other") %>%
  dplyr::group_by(Age, Variant) %>%
  summarise(Frequency = n()) %>%
  dplyr::ungroup() %>%
  na.omit() %>%
  ggplot(aes(x = Age, y = Frequency, group = Variant, fill = Variant)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_bw() +
  geom_text(aes(y = Frequency + 5, label = Frequency), 
            color = "gray20", position = position_dodge(0.9), size=3) +
  scale_fill_grey() +
  theme(legend.position = "top")
ggsave(here::here("images", "ge_age_stuff.png"), unit = "cm", width = 10)
```



## Bortuta

Prepare data

```{r}
citdata <- dstat %>%
  dplyr::rename(Texttype = texttype,
                Occurrence = nge,
                GeneralExtender = ge,
                Variant = ges) %>%
  dplyr::filter(Variant != "0") %>%
  dplyr::mutate(FileSpeaker = paste0(File, "$", Speaker, sep = "")) %>%
  dplyr::select(-Id, -Wordsunit, -GeneralExtender, -File, -Speaker, 
                -Occurrence, -ChildhoodCounty, -Occupation) %>%
  dplyr::mutate(Variant = droplevels.factor(Variant),
      Region = dplyr::case_when(Region == "NorthernIreland" ~ "NI",
                                Region == "RepublicOfIreland" ~ "RoI",
                                Region == "Interstate" ~ "Mixed",
                                TRUE ~ "NA")) %>%
  dplyr::mutate_if(is.character, factor) %>%
  dplyr::filter(Texttype == "Private")
# inspect
head(citdata)
```


```{r}
check <- citdata[complete.cases(citdata),]
nrow(citdata); nrow(check)

# checks
sum(table(citdata$Region))
sum(table(citdata$DateOfRecording)) # problem
sum(table(citdata$Gender))
sum(table(citdata$Age))
sum(table(citdata$ResidenceType)) # problem
sum(table(citdata$Education)) # problem
sum(table(citdata$ReligiousAffiliation)) # problem
sum(table(citdata$Texttype))
sum(table(citdata$AudienceSize))
sum(table(citdata$ConversationType))
sum(table(citdata$Variant))
sum(table(citdata$FileSpeaker))
```

```{r}
citdata <- citdata %>%
  dplyr::select(-DateOfRecording, -ResidenceType, -Education, 
                -ReligiousAffiliation, -Texttype) %>%
  na.omit(.)
# inspect
nrow(citdata) 
```


```{r}
# set.seed 
set.seed(2019120207)
# initial run
boruta1 <- Boruta(Variant ~., data=citdata)
print(boruta1)
```

```{r}
# extract decision
getConfirmedFormula(boruta1)
```


```{r}
plotImpHistory(boruta1)
```




## CIT


Perform CIT analysis


```{r}
set.seed(111)        # set.seed
# apply bonferroni correction (1 minus alpha multiplied by n of predictors)
control = ctree_control(mincriterion = 1-(.05*ncol(citdata)-1))
# create initial conditional inference tree model
citd.ctree <- partykit::ctree(Variant ~ Region + Gender + Age + AudienceSize,
                    data = citdata)
plot(citd.ctree, gp = gpar(fontsize = 8)) # plot final ctree
```

save plot

```{r}
png(here::here("images", "cit_base.png"), width = 1200)
plot(citd.ctree, gp = gpar(fontsize = 14)) 
dev.off()
```

pretty plot

```{r freq, warning=F, message=F}
# extract p-values
pvals <- unlist(nodeapply(citd.ctree, ids = nodeids(citd.ctree), function(n) info_node(n)$p.value))
pvals <- pvals[pvals <.05]
# plotting
pcit_freq <- ggparty(citd.ctree) +
  geom_edge(size = .5, col = "gray80") +
  geom_edge_label(size = 3) +
  geom_node_label(line_list = list(aes(label = splitvar, size = 5),
                                   aes(label = paste0("N=", nodesize, ", p", 
                                                 ifelse(pvals < .001, "<.001",
                                                   paste0("=", round(pvals, 3)))))),
                  line_gpar = list(list(size = 8),
                                   list(size = 8)), 
                  ids = "inner") +
  geom_node_label(aes(label = paste0("N = ", nodesize)),
                  ids = "terminal", 
                  nudge_y = 0.01, 
                  nudge_x = 0.01, 
                  size = 3) +
  geom_node_plot(gglist = list(
    geom_bar(aes(x = Variant, fill = Variant),
             position = position_dodge(), color = "black"),
    theme_minimal(),
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 90, size = 8),
          axis.text.y = element_text(size = 8)),
      scale_fill_manual(values = gray.colors(length(names(table(citdata$Variant)))),
                        guide = F),
      scale_y_continuous(breaks = seq(0, 100, 25),
                         limits = c(0, 100)),
    labs(x = "", y = ""),
      geom_text(aes(x = Variant, group = Variant,
                    label = stat(count)),
                stat = "count",
                size = 2, 
                position = position_dodge(0.9), vjust = -0.7)),
    shared_axis_labels = TRUE)
# save plot ggsave
ggsave(plot = pcit_freq, here::here("images", "cit_pfreq.png"), 
       unit = "cm", width = 20, height = 15)
# show plot
pcit_freq
```




```{r}
# extract p-values
pvals <- unlist(nodeapply(citd.ctree, ids = nodeids(citd.ctree), function(n) info_node(n)$p.value))
pvals <- pvals[pvals <.05]
# plotting
pcit_stack <- ggparty(citd.ctree) +
  geom_edge() +
  geom_edge_label() +
  geom_node_label(line_list = list(aes(label = splitvar),
                                   aes(label = paste0("N=", nodesize, ", p", 
                                                      ifelse(pvals < .001, "<.001", paste0("=", round(pvals, 3)))), 
                                       size = 10)),
                  line_gpar = list(list(size = 13), 
                                   list(size = 10)), 
                  ids = "inner") +
  geom_node_label(aes(label = paste0("N = ", nodesize)),
    ids = "terminal", nudge_y = -0.0, nudge_x = 0.01) +
  geom_node_plot(gglist = list(
    geom_bar(aes(x = "", fill = Variant),
             position = position_fill(), 
             color = "black"),
      theme_minimal(),
      scale_fill_manual(values = gray.colors(length(table(citdata$Variant)))),
      scale_y_continuous(breaks = c(0, 1)),
    xlab(""), 
    ylab("Probability"),
    geom_text(aes(x = "", group = Variant,
                  label = ""),
              stat = "count", position = position_fill(), vjust = 1.1)),
    shared_axis_labels = TRUE)
# save plot ggsave
ggsave(plot = pcit_stack, here::here("images", "cit_pfreq.png"), 
       unit = "cm", width = 20, height = 15)
# show plot
pcit_stack
```



## CFA

```{r genex0164, message=FALSE, warning=FALSE}
# prepare data
cfadata <- citdata %>%
  dplyr::group_by(Region, Gender, Variant) %>%
  dplyr::summarise(Frequency = n())
# define configurations
configs <- cfadata %>%
  dplyr::select(Region, Gender, Variant)
# define counts
counts <- cfadata$Frequency
# perform cfa
cfares <- cfa(configs, counts)
cfasig <- cfares$table %>%
  as.data.frame() %>%
  dplyr::filter(sig.z == T)
# save
write.table(cfasig, here::here("tables", "sigcfa.txt"), sep = "\t")
# inspect
cfasig
```


# Outro{-}

We have reached the end of part 1 of the analysis.

```{r outro}
sessionInfo()
```



